Blaming the Typeless: \\ Scalable, Human-Centric Python Localization

### 2018-08-25 paper/main.tex
paper

Some features, such as ``program crashes here'' and features
related to term size, are fairly direct. Others, such as ``Child3 is
statement'', a contextual, syntactic feature that was found to be very
important, merit additional explanation. In practice, we found it to
encode complex control structures, like large loops, where fixing the crash
often involved adding or removing whole statements from the loop.
Similarly, features that determine if the parent node
is a list or dictionary are relevant because of FIXME. In beginner-written
Python we find that defects and typos often in the immediate child position
of a list or dictionary term because of FIXME. Finally, we note the
importance of our \lt{Unknown} type feature for expressions that are never
evaluated in the given trace: we find that they are relevant because of
FIXME. \fixme{Ben (etc.) must actually fill in real explanations here.}

@@

\begin{figure}
\begin{tikzpicture}
  \begin{axis}[
      xlabel=Clustering threshold parameter,
      ylabel=\# of naturally occuring program clusters,
      xmax=1.25,
      legend style={at={(0.5,0.25)},anchor=north}
    %   ytick distance=1000,
    %   title=Linkage: single
    ]
    \foreach \language in {ocaml, python}
    {
      \addplot+[mark=none] table[x=threshold, y=\language_cluster_counts_single, col sep=comma]{cluster_counts.csv};
    }
\legend{NATE OCaml data, Our Python data}
\end{axis}
\end{tikzpicture}
\caption{Number of natural program clusters in each dataset as a function
of clustering parameter (informally, how ``strict'' or ``tight'' the
clustering is). \fixme{Talk about how you sample to be fair to the smaller
datasets.} At all points our dataset demonstrates more diversity than
that of previous work.}
\label{fig-diversity}
\end{figure}

@@

% The exact number of clusters returned
% depends on the value we choose for the algorithm's threshold parameter, so we
% checked all thresholds at .01 intervals, starting at 0 and ending at the point
% where everything collapses into a single cluster. As shown in
% \figref{diversity}, regardless of the choice of threshold, our data breaks into
% more clusters than the NATE data. This confirms that our data is more
% heterogeneous.

@@

\fixme{We need to say something about how we controlled for number, lest
reviewers think we have more clusters simply because we have hundreds of
thousands of datapoints.}

\figref{diversity} shows the results of clustering on our dataset and the
Nate dataset. At a high level, the structural similarity in the graphs
confirms the comparability of cluster counts between the two datasets.  For
all values of the inconsistency threshold, there are more clusters in our
Python dataset than in the baseline OCaml dataset. This suggests that our
dataset contains a more diverse set of programs. Note that we are not
claiming any novelty or advances in clustering accuracy in this
determination (indeed, scalability concerns limited us to coarser
approaches); instead, our claim is that even with simple clustering,
it is clear that our dataset contains a greater diversity of programs, even
when controlling for size, than were considered by previously-published
evaluations. We view it as an advantage of our algorithm that it can apply
to multiple different program categories.

@@

In the programming languages community, a large body of work has focused on
localizing faults and providing better error messages, typically through
the use of type information. SHErrLoc~\cite{SHErrLoc} identifies a set of
incompatible type constraints, then chooses the likely type error using the
Bayesian heuristic that the program is probably mostly correct. We rely on the
complementary heuristic that faults will share characteristics
with previously recorded faults, and apply this heuristic to features beyond
just type information. Mycroft~\cite{MYCROFT-type-errors} modifies existing
type inference algorithms to produce a ``correcting set'' for a program with
a type error. Mycroft assumes that the minimal such set is the most
desirable to the user, and has no way to rank multiple equally
small sets. We instead use machine learning to agree with human judgements.
NATE~\cite{learning-to-blame} uses machine learning on static and contextual
features, including static types, to localize type errors in OCaml
code. We employ a similar approach to localize Python faults, but we use
dynamic features as well as static and contextual ones, we handle a wide
variety of errors beyond type errors, and we evaluate our approach on a set of
programs both more heterogeneous and almost two orders of magnitude larger than
the NATE dataset.

>>>>

Some features, such as ``program crashes here'' and features
related to term size, are fairly direct. Others, such as ``Child3 is
statement'', a contextual, syntactic feature that was found to be very
important, merit additional explanation. In practice, we found it to
encode complex control structures, like large loops, where fixing the crash
often involved adding or removing whole statements from the loop.
In contrast, features that determine if the parent node
is a list or dictionary are relevant as strong negative features: we find that
fixing bugs in beginner-written
Python rarely requires changing the elements of a list or dictionary literal.
Finally, we note the
importance of our \lt{Unknown} type feature for expressions that are never
evaluated in the given trace: we find that they are relevant because of
FIXME. \fixme{Ben (etc.) must actually fill in real explanations here.}

@@

\begin{figure}
\begin{tikzpicture}
  \begin{axis}[
      xlabel=Clustering threshold parameter,
      ylabel=\# of naturally occuring program clusters,
      xmax=1.25,
      legend style={at={(0.5,0.25)},anchor=north}
    %   ytick distance=1000,
    %   title=Linkage: single
    ]
    \foreach \language in {ocaml, python}
    {
      \addplot+[mark=none] table[x=threshold, y=\language_cluster_counts_single, col sep=comma]{cluster_counts.csv};
    }
\legend{Nate OCaml data, Our Python data}
\end{axis}
\end{tikzpicture}
\caption{Number of natural program clusters in the Nate dataset and in a
random sample of our dataset of the same size, as a function
of clustering parameter (informally, how ``strict'' or ``tight'' the
clustering is). At all points our dataset demonstrates more diversity than
that of previous work.}
\label{fig-diversity}
\end{figure}

@@

% The exact number of clusters returned
% depends on the value we choose for the algorithm's threshold parameter, so we
% checked all thresholds at .01 intervals, starting at 0 and ending at the point
% where everything collapses into a single cluster. As shown in
% \figref{diversity}, regardless of the choice of threshold, our data breaks into
% more clusters than the Nate data. This confirms that our data is more
% heterogeneous.

@@

\figref{diversity} shows the results of clustering on our dataset and the
Nate dataset. Since our dataset is much larger than Nate's, we
used a random sample of our data of the same size as the Nate dataset.
At a high level, the structural similarity in the graphs
confirms the comparability of cluster counts between the two datasets.  For
all values of the inconsistency threshold, there are more clusters in our
Python dataset than in the baseline OCaml dataset. This suggests that our
dataset contains a more diverse set of programs. Note that we are not
claiming any novelty or advances in clustering accuracy in this
determination (indeed, scalability concerns limited us to coarser
approaches); instead, our claim is that even with simple clustering,
it is clear that our dataset contains a greater diversity of programs, even
when controlling for size, than were considered by previously-published
evaluations. We view it as an advantage of our algorithm that it can apply
to multiple different program categories.

@@

In the programming languages community, a large body of work has focused on
localizing faults and providing better error messages, typically through
the use of type information. SHErrLoc~\cite{SHErrLoc} identifies a set of
incompatible type constraints, then chooses the likely type error using the
Bayesian heuristic that the program is probably mostly correct. We rely on the
complementary heuristic that faults will share characteristics
with previously recorded faults, and apply this heuristic to features beyond
just type information. Mycroft~\cite{MYCROFT-type-errors} modifies existing
type inference algorithms to produce a ``correcting set'' for a program with
a type error. Mycroft assumes that the minimal such set is the most
desirable to the user, and has no way to rank multiple equally
small sets. We instead use machine learning to agree with human judgements.
Nate~\cite{learning-to-blame} uses machine learning on static and contextual
features, including static types, to localize type errors in OCaml
code. We employ a similar approach to localize Python faults, but we use
dynamic features as well as static and contextual ones, we handle a wide
variety of errors beyond type errors, and we evaluate our approach on a set of
programs both more heterogeneous and almost two orders of magnitude larger than
the Nate dataset.
